An Analysis of a Rising Trend in Dollar Loss in Fire Incidents With and Without Sprinkler Systems in Toronto over a decade (2011-2022)

### 2024-01-24 outputs/paper/paper.qmd
Minor edits to these 3 files.

cleaned_fire_data <- read_csv("../data/cleaned_data.csv", show_col_types = FALSE)
```
```{r}
#| echo: false
#| eval: true
#| label: tbl-cleaned
#| tbl-cap: "Cleaned Data showing the first 6 rows"
#| tbl-cap-location: bottom
#| warning: false

@@

@tbl-full shows the table of Dollar Loss over the years in the presence of proper sprinkler systems. As shown in @tbl-trend, the table presents three columns: year, dollar loss and dollar loss in millions, where the dollar loss in millions has been used to plot the bar graph in @fig-dollar-loss-both.
```{r}
#| echo: false
#| eval: true
#| label: tbl-full
#| tbl-cap: "Full Sprinkler System Present"
#| warning: false
#| message: false

@@

@tbl-none shows the table of Dollar Loss over the years when there were no sprinkler systems present and has the same columns as @tbl-full. This table, i.e. @tbl-none and @tbl-full have been used to plot @fig-dollar-loss-both
```{r}
#| echo: false
#| eval: true
#| label: tbl-none
#| tbl-cap: "No Sprinkler System Present"
#| warning: false
#| message: false

@@

y_limits <- c(0, 100)
plot1 <- ggplot(cleaned_fire_data_full, aes(x = `Year`, y = `Dollar Loss in Millions`)) + 
  geom_col(alpha = 1.5, fill = "red") + 
  ylim(y_limits) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_x_continuous(breaks = 2011:2022) +
  labs(fill = "With Sprinkler")
plot2 <- ggplot(cleaned_fire_data_none, aes(x = `Year`, y = `Dollar Loss in Millions`)) + 
  geom_col(alpha = 1.5, fill = "blue") + 
  ylim(y_limits) + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_continuous(breaks = 2011:2022) +
  labs(fill = "Without Sprinkler")
# I read on how to add two plots next to each other for comparison from ChatGPT 4, its referenced to in usage.txt.
grid.arrange(plot1, plot2, ncol=2)
```

@@

Let us now delve deeper into the possible reasons of such a trend. Firstly, as the population increases over the years there are more chances of there being fire incidents. As the city of Toronto gets more densely populated there is a higher chance or probability of there being a fire incident than when it was less densely populated. Secondly, the city of Toronto has evolved over the years by constructing skyscrapers, apartments, offices etc. for the residents to live and work in. As the number buildings in general increase around us, there will consequently be more chances of there being fire incidents than before. Finally, we can not ignore that with an improvement in infrastructure of the city, i.e. building better roadway systems or highways, there has been an influx of residents of surrounding areas to come to Toronto to work. This can also lead to there being more fire incidents such as unfortunate car or vehicle fires.

>>>>

cleaned_fire_data <- read_csv("../data/cleaned_data.csv", show_col_types = FALSE)
```
```{r}
#| echo: false
#| eval: true
#| label: tbl-cleaned
#| tbl-cap: "Cleaned Data showing Dollar loss over 2011-2022 with indication of Sprinkler System"
#| tbl-cap-location: bottom
#| warning: false

@@

@tbl-full shows the table of Dollar Loss over the years in the presence of proper sprinkler systems. As shown in @tbl-trend, the table presents three columns: year, dollar loss and dollar loss in millions, where the dollar loss in millions has been used to plot the bar graph in @fig-dollar-loss-both.
```{r}
#| echo: false
#| eval: true
#| label: tbl-full
#| tbl-cap: "Total Estimated Dollar Loss over 2011-2022 with Full Sprinkler System Present"
#| warning: false
#| message: false

@@

@tbl-none shows the table of Dollar Loss over the years when there were no sprinkler systems present and has the same columns as @tbl-full. This table, i.e. @tbl-none and @tbl-full have been used to plot @fig-dollar-loss-both
```{r}
#| echo: false
#| eval: true
#| label: tbl-none
#| tbl-cap: "Total Estimated Dollar Loss over 2011-2022 with No Sprinkler System Present"
#| warning: false
#| message: false

@@

y_limits <- c(0, 100)
plot1 <- ggplot(cleaned_fire_data_full, aes(x = `Year`, y = `Dollar Loss in Millions`)) + 
  geom_col(alpha = 1.5, fill = "red") + 
  ylim(y_limits) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_x_continuous(breaks = 2011:2022) +
  labs(title = "Full Sprinkler System")
plot2 <- ggplot(cleaned_fire_data_none, aes(x = `Year`, y = `Dollar Loss in Millions`)) + 
  geom_col(alpha = 1.5, fill = "blue") + 
  ylim(y_limits) + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_continuous(breaks = 2011:2022) +
  labs(title = "No Sprinkler System")
# I read on how to add two plots next to each other for comparison from ChatGPT 4, its referenced to in usage.txt.
grid.arrange(plot1, plot2, ncol=2)
```

@@

Let us now delve deeper into the possible reasons of such a trend. Firstly, i believe that as the population increases over the years there are more chances of there being fire incidents. As the city of Toronto gets more densely populated there is a higher chance or probability of there being a fire incident than when it was less densely populated. Secondly, the city of Toronto has evolved over the years by constructing skyscrapers, apartments, offices etc. for the residents to live and work in. As the number buildings in general increase around us, there will consequently be more chances of there being fire incidents than before. Finally, we can not ignore that with an improvement in infrastructure of the city, i.e. building better roadway systems or highways, there has been an influx of residents of surrounding areas to come to Toronto to work. This can also lead to there being more fire incidents such as unfortunate car or vehicle fires.
